Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2017-12-29 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Correction

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Brief Summary of Viscous Instabilities}\label{section:ViscousInstabilities}

Flow instabilities may be due to pressure gradient along the interface of fluids, stress forces acting in opposite directions, mobility differences that may exceed critical limits and heterogeneities of the domain. %\red{For example, during water flooding, efficiency is related with capillary number. The higher the capillary number the higher the recovery of hydrocarbons (LINK THESE SENTENCES ! HOW THIS NEW SENTENCE IS RELATED TO THE PREVIOUS -- 'Flow instabilities ... of the domain' ... AND HOW IS IT RELATED WITH THE FOLLOWING SENTENCES?)}. 
When the driving force of the injected fluid overcomes the capillary force, hydrodynamic instabilities may occur, resulting in the collapse of the interface between fluids and start forming fingers.

\medskip
In oil exploration, viscous instabilities (\ie fingering) can become more pronounced due to heterogeneities (\ie natural fractures, different permeability characteristics in different zones) of geological formations, therefore sweep efficiency may be reduced or early breakthrough may occur \citep{riaz_2004, tavassoli_2015}. During immiscible CO$_{2}$ flooding $\left(\text{\eg in CO}_{2}\text{ enhanced oil recovery, CO}_{2}\text{-EOR, processes}\right)$, density of supercritical CO$_{2}$ is usually lower than the \textit{in situ} hydrocarbons, thus viscous fingering and/or channelling may occur.

\medskip
As mentioned above viscous flow instabilities can be found across many research areas, disciplines and scales, from chemical and medical separation systems to large scale geological reservoir fields. \citet{muskat_1934} investigated the flow of a fluid injected between two infinite parallel plates and the impact of capillary number ($N_{c}$),

\begin{equation}
N_{c} = \frac{\mu U}{\gamma},
\label{eq:capillary_number}
\end{equation}
where $\mu$ is the viscosity of the fuid, $\gamma$ is the surface tension and $U$ is a velocity characteristic of the moving interface.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Brief Summary of Viscous Instabilities}\label{section:ViscousInstabilities}
The study of viscous flow instabilities (\ie fingering) is particularly important in oil exploration due to heterogeneities (\ie natural fractures, different permeability characteristics in different zones) of geological formations. In this application, volumetric sweep efficiency \red{(KOSTAS: MAY BE YOU SHOULD ADD ONE SENTENCE DEFINING/EXPLAINING ``SWEEP EFFICIENCY'' AND ``BREAKTHROUGH'' ... AND HOW THEY ARE RELATED.)} may be reduced and early breakthrough may occur \citep{riaz_2004, tavassoli_2015}, leading to low oil production. During immiscible CO$_{2}$ flooding $\left(\text{\eg in CO}_{2}\text{ enhanced oil recovery, CO}_{2}\text{-EOR, processes}\right)$, density of supercritical CO$_{2}$ is usually lower than hydrocarbons, thus viscous fingering and/or channelling may occur.

\medskip
Flow instabilities may be due to: pressure gradient along the interface of fluids; stress forces acting in opposite directions; mobility differences that may exceed critical limits and; heterogeneities of the domain. %\red{For example, during water flooding, efficiency is related with capillary number. The higher the capillary number the higher the recovery of hydrocarbons (LINK THESE SENTENCES ! HOW THIS NEW SENTENCE IS RELATED TO THE PREVIOUS -- 'Flow instabilities ... of the domain' ... AND HOW IS IT RELATED WITH THE FOLLOWING SENTENCES?)}. 
When the driving force of the injected fluid overcomes the capillary force, hydrodynamic instabilities may occur, resulting in the collapse of the interface between fluids and start forming fingers. \red{(KOSTAS: YOU STILL NEED TO REWRITE THIS PARAGRAPH AND LINK BOTH SENTENCES. THE SECOND SENTENCE (``When the driving'') IS A GOOD START FOR THE PARAGRAPH, ALTHOUGH YOU SHOULD DETERMINE WHAT IS THE ''DRIVING FORCE''  THAT YOU REFERRED TO.)}

\medskip
Viscous flow instabilities can be found across several disciplines and scales, from chemical separation processes to geological reservoir fluids. \citet{muskat_1934} investigated the flow of a fluid injected between two infinite parallel plates (Hele-Shaw cell) and the impact on capillary number ($N_{c}$),
\begin{equation}
N_{c} = \frac{\mu U}{\gamma},
\label{eq:capillary_number}
\end{equation}
where $\mu$ is the viscosity of the fuid, $\gamma$ is the surface tension and $U$ is a velocity characteristic of the moving interface.
